---
title: "manuscript-outputs"
author: "L. Naslund"
date: "2023-07-24"
output: html_document
---
# Methods
Site description table
```{r site area km2}
st_area(deans) %>% as.numeric() *1e-6
st_area(catfish) %>% as.numeric() *1e-6
st_area(sister3) %>% as.numeric() *1e-6
st_area(pick) %>% as.numeric() *1e-6

res_time

ysi_summary %>% pivot_wider(id_cols = sampling, names_from = name, values_from =  c(mean, sd))

ysi_6.20_7.12
```

# Results
Pathways P1: Summary of fluxes
```{r}
# range of CO2-eq fluxes
range(co2_eq_summary$total_co2_eq)

# pathways contributing the most CO2-eq flux on each sampling and the proportions of pathway contributions to total flux
co2_eq_summary

# degassing contributions
bh_conc <- 98.66104183 -0.36507263	
bh_q_0913 <- 4/c(270, 214, 245)
bh_degas <- bh_conc * mean(bh_q_0913) * 86400 * 1e-6 *16.04 *(1/as.numeric(st_area(pick)))
bh_degas
bh_degas*27
((bh_degas)/(co2_eq_summary %>% filter(site == "Pick") %>% pull(total_co2_eq))) *100


deans1_conc <- 0.58316848-0.31061833
deans_q_0817 <- 4/c(13.85,13.8,13.37)
deans_degas <- deans1_conc * mean(deans_q_0817) * 86400 * 1e-6 *16.04 *(1/as.numeric(st_area(deans)))
deans_degas
deans_degas * 27
(deans_degas / (co2_eq_summary %>% filter(site == "Pick") %>% pull(total_co2_eq)))*100
```

Pathways P2: Summary of DO during sampling
```{r summary sampling do, echo = F}
do %>% filter(location == "Top") %>% 
  mutate(site = if_else(site == "Pick", "Blue Herron", site)) %>% 
  filter(is.na(sampling)==F) %>% 
  group_by(site) %>% 
  summarize(min_do = min(DO_mgL), max_do = max(DO_mgL), mean_do = mean(DO_mgL))
```

Pathways P2: Range of pathways 
```{r}
range(summary_pathways$flux_round_diffusion_CO2)
range(summary_pathways$flux_round_diffusion_CH4)
range(summary_pathways$flux_round_ebullition_CH4)
```

Times P1: Environmental variables CH4 diffusion
```{r all CH4 models}
dredge(global.model = ch4_all)
summary(ch4_do)
r2_nakagawa(ch4_do)

summary(ch4_temp)
r2_nakagawa(ch4_temp)

summary(ch4_int)
```

Times P2: Environmental variables CO2 diffusion
```{r all CO2 models}
dredge(global.model = co2_all)

r2_nakagawa(co2_sdo)
summary(co2_sdo)

summary(co2_intercept)

r2_nakagawa(co2_temp)
summary(co2_temp)

r2_nakagawa(co2_temp_nbh)
summary(co2_temp_nbh)
```

Times P3: Effect of sunny v cloudy conditions
```{r sunny v cloudy}
summary(suncloud.mod.ch4)
summary(suncloud.mod.co2)
```

Locations P1: Spatial pattern of ebullition 
```{r sp eb}
summary(mod7)
r2_nakagawa(mod7)

# no Blue Herron
summary(mod7.d)
r2_nakagawa(mod7.d)
1-exp(mod7.d@beta)

# Moran's I for Blue Herron
diff_moran %>% filter(site == "Blue Herron", p.value < 0.05, gas == "CO2") %>% arrange(observed)
```

Simulations P1: Consequences of limited sampling locations
```{r}
space_sims_combos %>% mutate(perc = 100* (co2_eq_diff/calc_co2_eq) ) %>% summarize(range = range(co2_eq_diff), range_perc = range(perc))
```


Simulations P2: Consequences of limited sampling
```{r}
# how wrong are you sampling one times 
time_combos %>% filter(ord1 == ord2) %>% mutate(perc = 1 - (sim_co2_eq/calc_co2_eq)) %>% summarize(min = min(co2_eq_diff), max = max(co2_eq_diff), min_perc = min(perc), max_perc = max(perc))

# times that minimize cumulative CO2 diffusion estimation error for sites other than Blue Herron
time_combos %>% filter(site != "Blue Herron") %>% mutate(co2_diffusion_difference_abs = abs(co2_diffusion_difference)) %>% group_by(ord1, ord2) %>% summarize(misestimate_sum = sum(co2_diffusion_difference_abs)) %>% arrange(misestimate_sum) %>% mutate(time = if_else((ord1 %in% c(1:3) & ord2 %in% c(4:7)), "day_night", "not"))

# median and max CO2 error
time_combos %>% filter(site != "Blue Herron") %>% mutate(co2_diffusion_difference_abs = abs(co2_diffusion_difference)) %>% summarize(median = median(co2_diffusion_difference_abs), min = min(co2_diffusion_difference), max = max(co2_diffusion_difference))

# percent of flux max error
time_combos %>% filter(site != "Blue Herron") %>% arrange(co2_diffusion_difference) %>% slice(1) %>% mutate(perc = (co2_diffusion_difference)/flux_round_diffusion_CO2) %>% pull(perc)

# diel variation in Ch4 diffusion
date_to_sampling(diffusion) %>% filter(!(sampling %in% c("Blue Herron 9/13 - 9/14", NA))) %>% group_by(sampling) %>% summarize(range = max(flux_ch4_g_hr)- min(flux_ch4_g_hr))

#median and max ch4 error
time_combos %>% filter(site != "Blue Herron") %>% mutate(ch4_diffusion_difference_abs = abs(ch4_diffusion_difference)) %>% summarize(median = median(ch4_diffusion_difference_abs), min = min(ch4_diffusion_difference), max = max(ch4_diffusion_difference))

# percent of flux max error 
time_combos %>% filter(site != "Blue Herron")  %>% mutate(ch4_diffusion_difference_abs = abs(ch4_diffusion_difference)) %>% arrange(desc(ch4_diffusion_difference_abs)) %>% slice(1) %>% mutate(perc = (ch4_diffusion_difference)/flux_round_diffusion_CH4) %>% pull(perc)

# times that minimize cumulative CH4 diffusion estimation error for sites other than Blue Herron
time_combos %>% filter(site != "Blue Herron") %>% mutate(ch4_diffusion_difference_abs = abs(ch4_diffusion_difference)) %>% group_by(ord1, ord2) %>% summarize(misestimate_sum = sum(ch4_diffusion_difference_abs)) %>% arrange(misestimate_sum) %>% mutate(time = if_else((ord1 %in% c(1:3) & ord2 %in% c(4:7)), "day_night", "not"))
```

Simulations P3: Consequences of limited sampling
```{r}
# minimize error for CO2 diffusion
time_combos %>% filter(site == "Blue Herron") %>% mutate(co2_diffusion_difference_abs = abs(co2_diffusion_difference)) %>% arrange(co2_diffusion_difference_abs) %>% mutate(time = if_else((ord1 %in% c(1:3) & ord2 %in% c(4:7)), "day_night", "not")) %>% mutate(perc_co2_diff = 100*(co2_diffusion_difference/flux_round_diffusion_CO2))

# consequences of sampling at 10 AM
time_combos %>% filter(site == "Blue Herron", ord1 == 1, ord2 == 1) %>% mutate(co2_diffusion_difference_abs = abs(co2_diffusion_difference))  %>% mutate(perc = co2_diffusion_difference/flux_round_diffusion_CO2)

# minimize error for CH4 diffusion
time_combos %>% filter(site == "Blue Herron") %>% mutate(ch4_diffusion_difference_abs = abs(ch4_diffusion_difference)) %>% arrange(ch4_diffusion_difference_abs) %>% mutate(time = if_else((ord1 %in% c(1:3) & ord2 %in% c(4:7)), "day_night", "not")) %>% mutate(perc = 100*(ch4_diffusion_difference/flux_round_diffusion_CH4))

```


# Discussion

P1: Limited sampling consequences
```{r limited sampling consequences}
# how wrong are you without ebullition 
co2_eq_summary %>% select(site, date, ebullition_CH4, without_eb) %>% arrange(without_eb)

# how wrong are you sampling one times 
time_combos %>% filter(ord1 == ord2) %>% mutate(perc = 1 - (sim_co2_eq/calc_co2_eq)) %>% summarize(min = min(co2_eq_diff), max = max(co2_eq_diff), min_perc = min(perc), max_perc = max(perc))

# how wrong are you by sampling one to four locations
space_sims_combos  %>% mutate(perc = 1 - (sim_co2_eq/calc_co2_eq)) %>% summarize(min = min(co2_eq_diff), max = max(co2_eq_diff), min_perc = min(perc), max_perc = max(perc))

space_sims_combos  %>% mutate(perc = 1 - (sim_co2_eq/calc_co2_eq))
head(space_sims_combos)
```

P1: Wolffia reservoir vs other reservoirs
```{r}
# relative total co2 eq emissions
wolffia_co2_eq <- co2_eq_summary %>% filter(site == "Pick") %>% pull(total_co2_eq)
co2_eq_summary %>% mutate(compare_wolffia = wolffia_co2_eq/total_co2_eq)  %>% filter(site != "Pick") %>% summarize(avg_comparison = mean(compare_wolffia))

# percentage co2 eq emissions from CH4 emissions at Blue Herron
co2_eq_summary %>% filter(site == "Pick") %>% pull(prop_ch4)
```

Temporal and spatial patterns P1: Windspeed
```{r}
range(ws$ws_mean, na.rm =T)
mean(ws$ws_mean, na.rm = T)
```

Temporal and spatial patterns P2: Surface DO and CH4 diffusion
```{r}
r2_nakagawa(ch4_do)
```

Magnitude of emissions P1: Mean emissions
```{r}
summary_pathways %>% mutate(ch4 = flux_round_diffusion_CH4 + flux_round_ebullition_CH4) %>% summarize(avg_co2 = mean(flux_round_diffusion_CO2), avg_ch4 = mean(ch4))

summary_pathways %>% filter(site == "Pick") %>% select(site, flux_round_diffusion_CH4, flux_round_ebullition_CH4)
```

Magnitude of emissions P3: Estimating all ponds emissions
```{r total emission car comparison}
# kg CO2 eq if all reservoirs are green
clarke_area * 1e6 * green_avg * 1e-3

# kg co2 eq if none of the reservoirs are green
clarke_area * 1e6 * not_green_avg * 1e-3

# plug resulting kg of CO2-eq emissions into EPA calculator https://www.epa.gov/energy/greenhouse-gas-equivalencies-calculator#results
```


# Main Text Figures

Figure 1: Summary of contribution of different pathways to total CO2-eq flux
```{r pathways summary fig, echo = F}
png(filename = "../4-Figures/co2-eq-summary.png", width = 8, height = 5, units = "in", res = 300)
all_fluxes_pathways %>% 
  mutate(flux_mean_g_m2_day_co2_eq =  flux_mean_g_day_co2_eq/area,
         flux_2.5_g_m2_day_co2_eq= flux_2.5_g_day_co2_eq/area, 
         flux_97.5_g_m2_day_co2_eq= flux_97.5_g_day_co2_eq/area, 
         pathway_gas = paste(pathway, gas, sep = " ") %>% as.factor() %>% fct_relevel("diffusion CO2", after = 0), 
         site_date = paste(site, date, sep = " ")) %>% 
  ggplot(aes(site_date, flux_mean_g_m2_day_co2_eq, fill = pathway_gas)) + 
  geom_bar(position = position_dodge(), stat = "identity")+
  scale_x_discrete(labels = c("Catfish 09/06/22" = "Catfish 06 Sept", "Catfish 09/18/22" = "Catfish 18 Sept", "Deans 08/16/22" = "Deans 16 Aug", "Deans 08/30/22" = "Deans 30 Aug", "Pick 09/13/22" = "Blue Herron 13 Sept", "Sister3 08/22/22" = "Sister 22 Aug"))+
  geom_errorbar(aes(ymax=flux_97.5_g_m2_day_co2_eq, ymin=flux_2.5_g_m2_day_co2_eq), position = position_dodge())+
  scale_fill_manual(values = c("#BFDBC0","#94A7C5", "#3D5170"), labels = c(expression(diffusion~CO[2]), expression(diffusion~CH[4]), expression(ebullition~CH[4])))+
  ylab(expression(CO[2]-equivalent ~ flux ~ (g~m^-2~day^-1)))+
  theme_bw()+
  theme(legend.text.align = 0, axis.title.x = element_blank(), legend.title = element_blank(), axis.text.x = element_text(angle=45, hjust=1), axis.text = element_text(size = 14, color = "black"), axis.title = element_text(size = 16, color = "black"), legend.text = element_text(size = 14))
dev.off()
```

Figure 2: Diel patterns of emissions
```{r knit diel figs, echo = F}
png(filename = "../4-Figures/diel-patterns.png", width = 7.5, height = 11.5, units = "in", res = 300)
plot_grid(co2_diel, ch4_diel, eb_diel, ncol = 1, labels = "AUTO", label_size = 18, vjust = 1.3)
dev.off()
```

Figure 3: Iterative site number consequences
```{r fig 3}
png(filename = "../4-Figures/interate_space.png", width = 8, height = 12, units = "in", res = 300)
plot_grid(iterate_space, iterate_space_legend, ncol = 2, rel_widths = c(0.7, 0.3))
dev.off()
```

Figure 4: Heat map different sampling times
```{r fig 4}
png(filename = "../4-Figures/diffusion_heatmaps.png", width = 9, height = 11.5, units = "in", res = 300)
plot_grid(co2_diff_hm, ch4_diff_hm, ncol = 1, labels = "AUTO", label_size = 18)
dev.off()
```


# Supplement Figures
Figure S1: Inset map
```{r si 1}
plot_usmap(include = c("GA", "AL", "SC", "FL"))

athens <- data.frame(lon = -83.360351, lat = 33.884606)
athens_trans <- usmap_transform(athens)

png(filename = "../4-Figures/inset.png", width = 12, height = 5, units = "in", res = 300)
plot_usmap(include = c("GA", "AL", "SC", "FL")) + 
  geom_point(data = athens_trans, 
             aes(x = x, y = y), 
             color = "black",
             size = 3)
dev.off()
```

Figure S2: Sunny vs. cloudy day fluxes
```{r}
png(filename = "../4-Figures/sunny-cloudy.png", width = 12, height = 5, units = "in", res = 300)
plot_grid(sc_ch4 + theme(legend.position = "none"), sc_co2+ theme(legend.position = "none"), sc_legend, rel_widths = c(0.425,0.425, 0.15), nrow = 1)
dev.off()
```

Figure S3
```{r}
png(filename = "../4-Figures/daily_traces.png", width = 11, height = 12, units = "in", res = 300)
plot_grid(do_plot, lux_plot, ws_plot, temp_plot, par_plot, ncol = 2, labels = "AUTO", label_size = 18, align = "hv")
dev.off()
```

Figure S4: Two week DO trace
```{r do time series, echo = F}
anno <- data.frame(
  x1 = c(ymd_hms("2022-09-06 08:00:00", tz = "America/New_York"), 
         ymd_hms("2022-09-18 08:00:00", tz = "America/New_York"), 
         ymd_hms("2022-08-16 08:00:00", tz = "America/New_York"), 
         ymd_hms("2022-08-30 08:00:00", tz = "America/New_York"), 
         ymd_hms("2022-09-13 08:00:00", tz = "America/New_York"), 
         ymd_hms("2022-08-22 08:00:00", tz = "America/New_York") ),
  x2 = c(ymd_hms("2022-09-07 08:00:00", tz = "America/New_York"), 
         ymd_hms("2022-09-19 08:00:00", tz = "America/New_York"), 
         ymd_hms("2022-08-17 08:00:00", tz = "America/New_York"), 
         ymd_hms("2022-08-31 08:00:00", tz = "America/New_York"), 
         ymd_hms("2022-09-14 08:00:00", tz = "America/New_York"), 
         ymd_hms("2022-08-23 08:00:00", tz = "America/New_York")),
  y1 = c(rep(0,6)),
  y2 = c(rep(11,6)),
  site = c("Catfish", "Catfish", "Deans", "Deans", "Blue Herron", "Sister"))

do_crop <- do %>% filter(location == "Top") %>% 
  filter(Date_Time > ymd_hms("2022-08-01 08:00:00"), Date_Time < ymd_hms("2022-09-30 08:00:00"))

png(filename = "../4-Figures/do-all-summary.png", width = 8, height = 7, units = "in", res = 300)
ggplot() +
   geom_rect(data = anno, aes(xmin = x1, xmax = x2, ymin = y1, ymax = y2), fill = "#a3cdff", alpha = 0.5)+
  geom_line(data = do_crop, aes(Date_Time, DO_mgL))+
  theme_bw()+
  ylab("DO (mg/L)") + 
  theme(axis.title.x = element_blank(), axis.text = element_text(size = 14, color = "black"), 
        axis.title = element_text(size = 16, color = "black"), strip.text = element_text(size = 14,  color = "black"))+
  facet_grid(rows = vars(site))
dev.off()
```

Figure S5: Relationship between diffusive flux of CH4 and surface DO
```{r}
png(filename = "../4-Figures/do-ch4.png", width = 8, height = 4, units = "in", res = 300)
ch4_diff_do
dev.off()
```

Figure S6: Diel patterns of CO2 diffusion without Blue Herron
```{r co2 no bh}
png(filename = "../4-Figures/co2-diffusion-no-blue-herron.png", width = 9, height = 5, units = "in", res = 300)
co2_diel_no_bh
dev.off()
```

Figure S7: Relationship between diffusive flux of CO2 and surface DO
```{r}
png(filename = "../4-Figures/do-co2.png", width = 8, height = 4, units = "in", res = 300)
co2_diff_do
dev.off()
```


Figure S8: Relationship between diffusive flux of CO2 and water temperature
```{r}
png(filename = "../4-Figures/do-temp.png", width = 8, height = 4, units = "in", res = 300)
co2_diff_temp
dev.off()
```


Figure S9: Raw values map CO2
```{r}
png("../4-Figures/point_maps_co2.png", width = 12, height = 15, units = "in", res = 300)
tmap_arrange(pick_co2_13, catfish_co2_06, catfish_co2_18, deans_co2_16,  deans_co2_30, sister_co2_22, ncol = 2, outer.margins = 0.05)
dev.off()
```

Figure S10: Moran's I normal site-days
```{r}
png(filename = "../4-Figures/moran.png", width = 8, height = 5, units = "in", res = 300)
ggplot(diff_moran %>% mutate(pval = if_else(p.value > 0.001, if_else(p.value <0.05, paste("p =", as.character(signif(p.value, 2))), ""), "p <0.001"), uniqueid = paste(site, date, gas)), aes(trip, observed))+
  geom_point() + 
  scale_x_discrete(labels = c("10:00", "13:00", "16:00", "19:00", "21:00", "01:00", "04:00", "07:00"))+
  geom_text_repel(aes(label = pval), nudge_y = 0.01, size = 2.3, segment.alpha = 0) +
  labs(y= "Moran's I")+
  theme_bw()+
  theme(axis.title.x = element_blank(), 
        axis.text.x = element_text(angle = 45, hjust = 1, size = 12, color = "black"), 
        axis.text.y = element_text(hjust = 1, size = 12, color = "black"),
        axis.title = element_text(size = 16, color = "black")) +
  facet_wrap(~uniqueid)
dev.off()

diff_moran %>% mutate(uniqueid = paste(site, date, gas))%>% filter(uniqueid == "Blue Herron 13-14 Sept CO2")
```

Figure 10a: Morans I for the days when the analyzer had a hole
```{r}
png(filename = "../4-Figures/broken-analyzer-moran.png", width = 5, height = 4, units = "in", res = 300)
ggplot(broken_morans %>% mutate(pval = if_else(p.value > 0.001, if_else(p.value > 0.05, "", paste0("p = ", as.character(signif(p.value, 2)))), "p < 0.001")), aes(trip, observed))+
  geom_point() + 
  scale_x_discrete(labels = c("10:00", "13:00", "16:00", "19:00", "21:00", "01:00", "04:00", "07:00", "11:00", "15:00", "20:00", "01:00", "06:00"))+
  geom_text_repel(aes(label = pval), nudge_y = 0.01, size = 3, segment.alpha = 0) +
  labs(y= "Moran's I")+
  #ylim(0, 0.45)+
  theme_bw()+
  theme(axis.title.x = element_blank(), 
        axis.text.x = element_text(angle = 45, hjust = 1, size = 12, color = "black"), 
        axis.text.y = element_text(hjust = 1, size = 12, color = "black"),
        axis.title = element_text(size = 16, color = "black")) +
  facet_wrap(~uniqueid, scales = "free_x")
dev.off()
```

Figure S11: Heatmap for times without Blue Herron
```{r}
png(filename = "../4-Figures/nbh_heatmaps.png", width = 13, height = 9.5, units = "in", res = 300)
plot_grid(co2_diff_nbh_hm, ch4_diff_nbh_hm, ch4_eb_hm_nbh, co2_eq_diff_nbh_hm, ncol = 2, labels = "AUTO", label_size = 18)
dev.off()
```

Figure S12: Heatmap for ebullition and total CO2-eq
```{r}
png(filename = "../4-Figures/eb_co2_eq_heatmaps.png", width = 9, height = 11.5, units = "in", res = 300)
plot_grid(ch4_eb_hm, co2_eq_hm, ncol = 1, labels = "AUTO", label_size = 18)
dev.off()
```


# Supplement tables

Table S1: Dissolved gas concentrations
```{r}
# The dissolved gas samples from Catfish on 9/18/22-9/19/22 were not handled properly, thus those values are excluded from the summaries.
gc_summarized %>% 
  filter(mdy(Collection.Date)>mdy("08/15/22"), Collection.Date!="9/19/2022") %>%
  arrange(Location) %>% 
  select(Site, Collection.Date, Location, CH4_avg_umol, CH4_sd, CO2_avg_umol, CO2_sd) %>% 
  mutate(CH4_avg_umol = round(CH4_avg_umol, 3), CH4_sd = round(CH4_sd, 3), CO2_avg_umol = round(CO2_avg_umol,3), CO2_sd = round(CO2_sd, 3))
```

Table S2: Flux estimates
```{r}
all_fluxes_pathways  %>% select(site, date, pathway, gas, flux_mean_g_m2_day:flux_97.5_g_m2_day)  
```

Table S3: Environmental variables model estimates
```{r}
dredge(global.model = co2_all)

r2_nakagawa(co2_sdo)
summary(co2_sdo)

summary(co2_intercept)

r2_nakagawa(co2_temp)
summary(co2_temp)

summary(co2_temp_do_top)
r2_nakagawa(co2_temp_do_top)

summary(co2_do_do)
r2_nakagawa(co2_do_do)

dredge(global.model = ch4_all)
summary(ch4_do)
r2_nakagawa(ch4_do)

summary(ch4_temp)
r2_nakagawa(ch4_temp)

summary(ch4_int)
```

```

